### 2020-04-09 documentation/systemModel/formalized_work.tex
recursion and secrecy capacity fixed

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Solution Outline}
\author{Peter Hartig}

\begin{document}
\maketitle
%\tableofcontents
%\section{Solution Outline} 
\subsection{Model}

Using the model from the pilot decontamination work we say the total channel is described by 
\begin{equation}
\mathbf{D} = \sum_{k=1}^{K}\mathbf{\mathbf{D}}_k 
\end{equation}
We now consider how to ultimately arrive at an expression for the AED of the covariance matrix $\mathbf{D}\mathbf{D}^H$.
\subsection{Approach 1} 
The expanded covariance matrix $\mathbf{D}\mathbf{D}^H$ contains both Hermetian (e.g. $\mathbf{D}_k \mathbf{D}_k ^H$) and non-hermetian terms ($\mathbf{D}_k \mathbf{D}_i^H, i \neq k$). Noting that the use of multiplicative free convolution and additive free convolution both require that the involved operators are \emph{hermetian} we could attempt to find the R-Transforms of the terms $\mathbf{D}_k \mathbf{D}_k ^H$ and $(\mathbf{D}_k \mathbf{D}_i^H+\mathbf{D}_i \mathbf{D}_k^H)$ which are both self-adjoint. 
For the terms $\mathbf{D}_k \mathbf{D}_k ^H$, we can find the S-Tranform using the methods shown in \ref{Polynomial_terms} and \cite{muller2002asymptotic}. From this S-Transform, we can move to the R-Transform in order to use additive free convolution. Finding the R-Transform of the terms $(\mathbf{D}_k \mathbf{D}_i^H+\mathbf{D}_i \mathbf{D}_k^H)$ is, however, not clear because the components of this sum are not are not self-adjoint.
Instead we consider a second approach.

\subsection{Approach 2} 
This approach is largely based on the ideas presented in \cite{muller2012channel} and which are applied in \cite[Eqns. 54-56]{muller2014blind}. The exampled \cite{muller2012channel} uses a very similar model to the IRS work.  The idea behind this approach is based on the property of the Stieltjes transform 
	\begin{equation}\label{ST_property}
	G_{\lambda^2}(s) = \frac{G_{\lambda}(\sqrt{s})- G_{\lambda}(-\sqrt{s})}{2\sqrt{s}}.
	\end{equation} (*TO DISCUSS Can we find an equivalent to this property in the S-transform domain?)
	Note that this property also requies that the matrix $D$ is bi-unitarily invariant.
	Rather than looking for the Stieltjes transform of the eigenvalues of $\mathbf{D} \mathbf{D}^H$ we will
	first find the Stieltjes transform of the singular values of $\mathbf{D}$. 
	Given the definition of the singular values $\sigma(\mathbf{D}) = \sqrt{\lambda(\mathbf{D} \mathbf{D}^H)}$ and 
	noting that 
	for a covariance matrix (positive semi-definite) all eigenvalues are non-negative, the eigen
	values of $\mathbf{D} \mathbf{D}^H$ are the squared singular values of  $\mathbf{D}$. Therefore, if 
	we can find $G_{\mathbf{D}}(s)$, we can use Property \ref{ST_property} to find $G_{\mathbf{D} \mathbf{D}^H}(s)$.
	 This is performed using the following steps.

\begin{enumerate}

\item 
	We first find the Stieltjes transform, $G_{\mathbf{D}_{k}\mathbf{D}_{k}^H}(s)$, for the AED of the 
	covariance matrix of individual terms, $\mathbf{D}_{k}$.
	This is discussed further in Section \ref{Polynomial_terms} and \cite[Section II]{muller2002asymptotic}.
\item
	Using Property \ref{ST_property} we find the Stieltjes transform of the symmetrized 
	singular value decomposition of $\mathbf{D}_{k}$, $\tilde{G}_{\mathbf{D}_{k}}(s)$. The symmetrized AED defined 
	in \cite[Eqn. 38]{muller2012channel} is given by 
	\begin{equation}
	\tilde{p}(x) = \frac{p(x)+P(-x)}{2}
	\end{equation}
\item
	Taking the R-transform, $\tilde{R}_{\mathbf{D}_{k}}(w)$ of $\tilde{G}_{\mathbf{D}_{k}}(s)$, we then use the free additive convolution property in order to 
	find the R-transform of the entire equivalent channel $\tilde{R}_{D}(w) = \sum_{k=1}^{K} \tilde{R}_{\mathbf{D}_{k}}(w) $ . 
	* Note that while additive free convolution holds for $\tilde{R}_{\mathbf{D}_{k}}(w)$, multiplicative free convolution does
	 not.
\item
	Using $\tilde{R}_{D}(w)$ we then find $\tilde{G}_{D}(s)$. We then use property \eqref{ST_property} in the reverse
	 order to get $\tilde{G}_{DD^H}(s)$  which allows us to evaluate the  symmetrized AED.

\item 
	Because all eigenvalues are are non-negative for the covariance matrix $\mathbf{D}\mathbf{D}^H$
	the symmetrized AED is equal to the true AED when evaluated at non-negative values and can be divided by 2
	when evaluated at 0.

\end{enumerate}

\subsection{Solving for $G_{\mathbf{D}_{k}\mathbf{D}_{k}^H}(s)$}\label{Polynomial_terms}
Using the model from the IRS work, the individual terms of the channel are given by
\begin{equation}
\mathbf{D}_{k} = G_B \boldsymbol{\Theta}_N H_N \cdots \boldsymbol{\Theta}_1 H_1
\end{equation}
and 
\begin{equation}
C_N = \mathbf{D}_{k}\mathbf{D}_{k}^H = [G_B \boldsymbol{\Theta}_N H_N \cdots \boldsymbol{\Theta}_1 H_1]
[H_1 \boldsymbol{\Theta}_1 \cdots H_N \boldsymbol{\Theta}_N G_B]
\end{equation}
\begin{itemize}
\item 
	Because commuting matrix multiplication does not change the non-zero eigenvalue distribution, we consider 
	the matrix 
\begin{equation}
\tilde{C}_N = [\boldsymbol{\Theta}_1^H \cdots H_N^H \boldsymbol{\Theta}_N^H G_B^H][G_B \boldsymbol{\Theta}_N H_N \cdots \boldsymbol{\Theta}_1 H_1 H_1^H]
\end{equation}

\item 
	We use the rotational property of the Stieltjes transform and S-Transform. 
\begin{equation}
S_{C_N}(z) = \frac{z+1}{z+\chi_N} S_{\tilde{C}_N}(\frac{z}{\chi_N}).
\end{equation}
For the two terms of $\tilde{C}_N$, 
$ H_1 H_1^H$ and $[\boldsymbol{\Theta}_1^H \cdots H_N^H \boldsymbol{\Theta}_N^H G_B^H][G_B \boldsymbol{\Theta}_N H_N \cdots \boldsymbol{\Theta}_1]$
we can use multiplicative free convolution .

\item
	For the gaussian IID case, $ H_1 H_1^H$  has the marcenko-pastur distribution which has a well defined S-Transform.

\item 
	Iterating the rotation property with the second term 
	\begin{equation}
	[\boldsymbol{\Theta}_1^H H_2 \cdots H_N^H \boldsymbol{\Theta}_N^H G_B^H]
	[G_B \boldsymbol{\Theta}_N H_N  \cdots H_2^H \boldsymbol{\Theta}_1]
	\end{equation}

	\begin{equation}
	[ \boldsymbol{\Theta}_N H_N  \cdots H_2^H \boldsymbol{\Theta}_1]
	[\boldsymbol{\Theta}_1^H H_2 \cdots H_N^H \boldsymbol{\Theta}_N^H G_B^H G_B]
	\end{equation}

	 it is seen that ultimately, the phase matrices will cancel out and the result is a product of well known S-Transforms.

\item 
*TO DISCUSS If there is no equivalent to the Property \ref{ST_property} in the S-Transform domain, how do we 
find a closed form of the Stieltjes transform given an arbitrary S-transform?
\end{itemize}

\bibliography{bibliography}
\end{document}