### 2020-07-02 documentation/Meeting Notes/07_01_20.tex
Added function and script for projection results. TODO noralize power with projection

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\maketitle
\tableofcontents

\section{Numerical Results}
First I confirm that the methods I use to obtain the AED from a known, closed-form S-Transform works.
The numerical results  used from comparison are generated by averaging 100 realizations of the system for size $N=100 \times 100$.
\begin{itemize}
\item
	Channel 
	\begin{equation}
	\mH  \sim \CNO
	\end{equation}
	For equal power allocation 
	\\
	Numeric Capacity = $  1.4286$ 
	\\
	Asymptotic Capacity = $ 1.4129 $

\item
	Channel 
	\begin{equation}
	\mH \phase \mH  
	\end{equation}
	with $\mH \sim \CNO$.
	For equal power allocation 
	\\
	Numeric Capacity = $ 1.4227$ 
	\\
	Asymptotic Capacity = $ 1.4060 $

\item
	Channel 
	\begin{equation}
	\mC\mH  
	\end{equation}
	with $\mC$ as the exponential correlation matrix. 
	For equal power allocation the capacities are plotted below. Because the s-transform for the correlation has a quadratic form, I show the solutions for both the $+/-$ forms of the sign.

	\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/correlation_comparison}
	  \caption{Equal power capacity with correlation form 1.}
	\end{figure}
	\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/correlation_comparison2}
	  \caption{Equal power capacity with correlation form 2.}
\end{figure}
\item
	These are the steps I use to move from the closed-form S-Transform of the channel to the AED.
	The S-Tranform of the exponential correlation model is derived in \cite{skupch2005free}
	\begin{enumerate}
	\item
		Beginning with the S-Transform, $S(z)$, I find $\gamma(s)$ using the fixed-point equation
		\begin{equation}
			S(\gamma(z))\gamma(z) - z - z\gamma(z) = 0 
		\end{equation}
	\item
		I find the Stieltjes transform using
		\begin{equation}
			G(s) = \frac{1}{s} (1+\gamma(\frac{1}{s}))
		\end{equation}			
	\item 
		I recover the AED using 
\begin{equation}
p(x) = \frac{1}{\pi} \underset{y \rightarrow \infty}{\text{lim}} \; \text{Img}(G(x+jy))
\end{equation}
	\end{enumerate}

\item
	I checked the equation provided for the gamma transform of the the exponential correlation model and observed the following AEDs.
		\begin{figure}[H]
	\includegraphics[width= 15cm,height = 10cm]{figures/exponential_correlation_aed}
	  \caption{Comparison of AED for the exponential correlation model using results from \cite{skupch2005free}}
	\end{figure}

\end{itemize}
\section{Questions}
\bibliography{bibliography}

\end{document}