### 2020-07-14 documentation/Paper/abstract.tex
Checked secrecy capacity for correlated. As expected secrecy capacity increases with correlation for the IRS channel.

This work investigates the impact of passive reflective surfaces on the capacity of MIMO communication systems. This using the tools of free probability, this system is analyzed in the asymptotic domain to derive an expression for the system capacity. 
A system is considered first in the case of fully uncorrelated channels. In this case it is shown that the passive control of such a reflective surface has no impact on the capacity of the system. The system is then generalized to allow for correlation within the channel. In this case is it shown that the selection of the phases at the IRS still has an impact on the system. Finally we compare the capacity of the correlated channel with optimized phases at the IRS to the uncorrelated channel to show that....?

This work provides an upper bound on the capacity of the RIS assisted MIMO system. This upper bound is a useful tool with which to analyze the performance of the various sub-optimal optimization approaches used. 
Question: would the capacity of the finite , phase-optimized system with correlation ever exceed that of the asymptotic case?