### 2020-07-21 documentation/Paper/information_theory_background.tex

A key metric of interest for the channel described in the previous section is the mutual information between the transmitted and received signal. If the channel $\mHt$ is known at the receiver, the mutual information is given by 
\begin{equation}\label{mut_ent}
I(\vy;\vx) = h(\vy) - h(\vy|\vx).
\end{equation}
If both $\vx$ and $\vn$ are circularly symmetric Gaussian random vectors, $\vy$ is also circularly symmetric Gaussian with entropy given by \cite{telatar1999capacity} 
\begin{equation}\label{entropy}
h(\vy) = \Expect[\Log(|\pi e \mathbf{Q}_{y}|)]
\end{equation}
 for $\mathbf{Q}_{y} = \Expect{[\vy \vy^H]}$. Note also that for a given $\mathbf{Q}_{y}$ the entropy is maximized if and only if $\vy$ is circularly symmetric Gaussian.

For a given transmit covariance matrix, $\mathbf{Q_x}$ and $\Expect\left[\vn\vn^H \right] = \mI$, substituting equation \eqref{entropy} into equation
\eqref{mut_ent} and simplifying gives
\begin{equation}\label{mut_inf}
\Expect \left[\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)\right].
\end{equation}
Using the model for $\mHt$ from the previous section, and maximizing with respect to both the
transmit covariance matrix and phases at the IRS gives an expression for the system capacity

\begin{equation}\label{capacity}
\mathbb{C} = \underset{\boldsymbol{\Phi},\mathbf{Q_x}}{\mathop{max}} \Expect \left[\Log(|\mathbf{I} + \mHt \mQ_x \mHt^H|)\right]
\end{equation}
or equivalently,
\begin{equation}\label{capacity_tricky}
\mathbb{C} = \underset{\boldsymbol{\Phi},\mathbf{Q_x}}{\mathop{max}} \Expect \left[\sum_{i=1}^{N}\Log(\mathbf{I} + \lambda_i)\right].
\end{equation}
where $\lambda_i$ are the eigenvalues of $\mHt \mQ_x \mHt^H$.

Recalling that in general we are interested in evaluating the capacity in order to determine a feasible transmission rate for a given communication channel, equation \eqref{capacity_tricky} poses a couple of hurdles to evaluation. First, we need to be able to evaluate an expectation over the joint pdf $p(\lambda_1 \cdots	 \lambda_N)$. Second, there is an optimization problem whose solution will generally depend on the distribution of the channel which would require a closed-form of the optimization problem solution. 
Two simplifications will be used in order to move forward. 
First, we will enforce that the transmitter chooses a deterministic $\mQ_x$. 
By enforcing a deterministic choice of $\mQ_x$, we no longer have to consider this component of the optimization.
Second we will normalize the channel, and therefore the received power, to allow the received signal covariance matrix eigenvalue distribution to converge.

If $\mHt$ is not known at the transmitter, a common choice for the transmitter signal covariance matrix is  $\mQ_x = \frac{P_{\text{total}}}{N_t}\mathbf{I}$, giving the simplified form of equation \eqref{capacity}
\begin{equation}\label{no_csi_capacity}
\mathbb{C} = \underset{\phase}{\mathop{max}} \; N_R \Expect[\Log(1 + \frac{P_{\text{total}}}{\sigma_nN_t}\lambda_i)].
\end{equation}
where $\lambda_i$ are the eigenvalues of $\mHt \mHt^H$.
In this form, the linear scaling of capacity with N is clear.
If we assume that $\Expect\left[ \vy \vy^H \right]$ has a converging eigenvalue distribution as the system dimensions increase to infinity, we can expand the expectation to get the expression
\begin{equation}\label{no_csi_capacity_aed}
\mathbb{C} = \underset{\phase}{\mathop{max}} \; N_R  \int_{x=0}^{\infty}\Log(1 + \frac{1}{\sigma_n}\lambda)p_{\lambda\lambda^H}(x) dx
\end{equation}
in which the bounds of the integral reflect the range for the eigenvalues of a positive semidefinite matrix. In the following we will refer to this distribution as the Asymptotic Eigenvalue Distribution (AED).

From this, we see that if the normalized channel has a converging eigenvalue distribution we can simply scale the capacity for a single channel by the system size $N_R$.

Note that because the choice of $h(\vx) $ and $h(\vn)$ maximize $h(\vy)$, equation \eqref{capacity} provides an upper bound on the capacity of the channel for any choices of the distribution of $\vx $ with Gaussian noise $\vn$. (Perhaps mention that normally we cannot generate continuous Gaussian x so we use channel coding).

Note that because $\Log$ is a concave function, we can upper bound the mutual information with
\begin{equation}\label{mut_inf}
\Expect[\Log(1 + \frac{1}{\sigma_n}\lambda)] \leq \Log(1 + \frac{1}{\sigma_n}\Expect[\lambda])
\end{equation}