### 2020-07-21 documentation/Paper/irs_optimization.tex

The previous section looked at a select few entries from the substantial literature investigating the capacity of MIMO systems.
This analysis can often be difficult because the definition of capacity implies an optimization over any controlled variables. 
While in select cases this optimization may yield a closed form solution that may be used in the analysis, this is not generally the case. 
In fact, the current literature on RIS system has skipped this analysis entirely because the RIS optimization is a non-convex problem. Rather 
numerous algorithms have been proposed and analyzed numerically. We now briefly review a few selected results from the current RIS literature. 
\par
Go through and review
\par
One key point to note is that in none of the above works and indeed so far as the author's knowledge have these works considered relevant case of a MIMO channel with channel correlation.