### 2020-07-23 documentation/Paper/correlation_review.tex

\par
Early investigation which lead to the current interest in MIMO systems showed the key property that with constant transmit power, the capacity (and mutual information) in an $N \times N$ MIMO system with independent, i.i.d Gaussian channels scales linearly in $N$ \cite{foschini1998limits}. While the assumption of \emph{identically distributed} channels is reasonable when adding more antennas, the assumption of \emph{independent} channels between each pair of transmit and receive antennas is not. To understand the potential for leveraging this capacity scaling in deployed massive MIMO systems, the investigation of correlated MIMO channels has gained substantial interest. In the following, a number of these studies are reviewed. Some of these chosen works use channel correlation models similar to the one we will develop for the RIS system model. In particular, this method models the correlation between antennas as decreasing  exponentially with the distance between antennas. Definitely need to show this is a valid model (Cite).
\par
Might need to give more motivation for the use of asymptotic analysis.
\par
One method that has been shown to reduce the complexity of this calculation is to allow the system size $N \rightarrow \infty$ while normalizing the received power. 
This is motivated by the observation that when the random channel is properly normalized, the eigenvalues of the channel covariance matrix often converge to a deterministic distribution which will be referred to as the Asymptotic Eigenvalue Distribution (AED). 
Because the received power would generally not be normalized as the system size in increased, results found while using this normalization be extended by simply scaling the asymptotic result with $N$.
(Still need to clarify the types of MIMO gains that are observed under this normalization and how to deal with non-zero mean components).
\par
Evaluating the capacity for the case in which the transmitter has channel state information (CSIT ) can be difficult as the solution to this optimization is not generally in a closed-form. In \cite{loyka2001channel}, this problem is avoided using an upper bound on the capacity of a general MIMO channel derived from Jensen's inequality. Using this upper bound and the exponential correlation model, an interpretation is given in which increasing correlation between antennas corresponds to a decrease in the average SNR at the receiver. 
Similarly, \cite{chuah2002capacity} investigates the exponentially correlated MIMO system using close approximations to compares optimal versus equal power distribution at the transmitter for cases with high and low SNR. One interesting finding from this work is that for the case with low
SNR (what does this mean?) the water-filling capacity is actually increased with correlation while in all other cases, correlation has a negative impact on the capacity. The intuition because this finding is that the equal power case is only impacted by the SNR and not the distribution of the channel eigenvalues (unlike the water-filling case).
Finally, \cite{skupch2005free} used multiplicative free convolution from free probability theory to analyze the exponential correlation model at either the transmitting or receiving antenna. Similarly, they also observe a decrease in the capacity as correlation increase but only at high values of correlation in the system. Part of this result will be used in our analysis. 
Finally,
\cite{taricco2008asymptotic} analyzed the correlated Ricean fading channel using the Replica Method. 
In general it is observed that the slope of the linear scaling of (mutual information) is reduced by the addition of correlation. 
\par
Before moving on to apply any of these asymptomatic results with our own system model, it is fitting to consider two questions. 
First, whether or not the assumption of  $N \rightarrow \infty$ is reasonable? Intuitively, we would expect that under correlation this convergence may occur only at very high $N$. This result was confirmed in  \cite{martin2004asymptotic} using the exponential correlation model at both the transmitter and receiver. 
Second, what values for the parameters of the correlation model are reasonable for a real-world system?
In test data collected with channel parameters (carrier frequency, bandwidth, power...) relevant to 3G and 4G, wireless correlation between channels was shown to be as high as $0.5$ for a point-to-point, $4 \times 4$ (linear array?) system \cite{martin2000multiple} (TODO justify if this is a comparison for current systems).