### 2020-07-23 documentation/Paper/paper.tex

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{float}
\input defs.tex
\bibliographystyle{alpha}
\graphicspath{ {./figures/} }
%\hyphenpenalty=10000

\title{Free Probability Analysis for Reflect Surface aided Massive MIMO systems}
\author{Peter Hartig}

\begin{document}
\maketitle
\begin{abstract}
\input{abstract}
\end{abstract}
%
%\newpage
\tableofcontents
\chapter{Introduction}
\section{Background}\label{Background}
\input{background_opening}
\par
\subsection{MIMO with correlation}\label{mimo_corr}
\input{correlation_review}
\par
\subsection{RIS Optimization}\label{irs_opt}
\input{irs_optimization}
\par
In summary, this section has highlight two primary points. First, that channel correlation can degrade the capacity scaling of massive MIMO systems. Second, that an RIS may provide additional degrees of freedom with which to increase capacity. We now investigate these questions be formulating and analyzing and RIS-aided, MIMO system with channel correlation.

\section{Notation}
\input{notation}
\section{Outline}
The following sections are now previewed. Section \ref{system_model} covers the system model used in the remainder as well as a review of the necessary theory to analyze the resulting channel. In section \ref{Results}, the theoretical results and corresponding numerical results are discussed.

\chapter{Analyzing the RIS System}\label{system_model}
\section{Analysis Preliminaries}
\subsection{Power Normalization}
\input{power_normalization}
\subsection{Mutual Information}\label{sectiond:mut_info}
\input{information_theory_background}
\subsection{Free Probability}\label{Free_Prob_Intro}
\input{free_probability}
\section{RIS Analysis}
\subsection{The Asyptotic IRS Channel}
\input{irs_channel}
\subsection{Non-Asymptotic Results}
\input{non_asymptotic}
\chapter{Results}\label{Results}
\section{Asymptotic Results}

\subsection{Asymptotic Results Discussion}

We first consider the case in which all matrices of the channel given by (ref) are free. As shown in (Appendix), in this case the 
phase shifts at the RIS, $\phase$, cancels out of the AED expression. 
By considering the case in which a single transmitter and receiver communicate over the channel $\mathbf{h}_r^T \phase \mathbf{h}_t$ it is clear that
the impact of the phase matrix must diminish under the assumptions of free probability, \ie  as the number of antennas at transmitter and receiver increase asymptotically and that the component matrices of the channel are free. 
\par
While the realized capacity of the asymptotic channel converge towards the asymptotic capacity as $nt$ and $nr$ increase, the rate of this convergence is particularly relevant in this application as this implies that the number of rf-chains at the transmitter and receiver approach infinity. One notable observation is that this rate
decreases as the correlation in the channel increase. Under these circumstances, it may still be beneficial to perform some optimization of the phases at the IRS. 
\begin{itemize}
\item

\end{itemize}
\subsection{Asymptotic Assumptions Under Correlation}
First mention that for iid random matrices, convergence is very fast (numerical). Then mention show for correlation within matrix.

\chapter{Conclusion}
\chapter{Appendix}
\input{appendix}
\bibliography{bibliography}
\end{document}