### 2020-07-27 documentation/Optimization Work/07_24_20.tex
Numerial evidence for phase optimization at high correlations

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\maketitle
\tableofcontents

\section{Numerical Results}
I use the system model given by 
\begin{equation}
\vy = \mH_{Total} \mW \vs  + \vn
\end{equation}
with

	\begin{equation}
	\mathbf{H}_{Total} = \mathbf{R}_{R}^{\frac{1}{2}}(\underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{R}_{S}^{\frac{1}{2}}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}})\mathbf{R}_{T}^{\frac{1}{2}}.
	\end{equation}

Using a matched filter  and precoder which satisfies the transmission power constraint gives (normalize each column)
\begin{equation}
\mW = \mathbf{H}_{Total}^H
\end{equation}

Numerical results for $s_k \in \mathcal{NC}(0,1)$, $n_k \in \mathcal{NC}(0,1)$
\begin{itemize}
\item
zf mu: 202.0989509146119
zf std: 57.23864582855936
\item
matched mu: 476.69650954785794
matched std: 239.3529067968275
\end{itemize}

\bibliography{bibliography}

\end{document}