### 2020-08-07 documentation/Paper/abstract.tex

Reflective surfaces with controlled reflected phase (IRS) have been the subject of interest in recent years in order to control the wireless propagation environment. This work investigates the impact of an IRS on the capacity of MIMO communication systems. Using the tools of free probability, a system is analyzed in the asymptotic domain to derive an expression for the system capacity. A key component of this analysis is considering the correlation common to massive MIMO systems, an aspect of the MIMO system not yet considered in previous work on IRS assisted systems. A key finding is that the choice for the phase of reflected waves has no impact for the asymptotic analysis. Motivated the contradiction between this finding and the substantial previous literature investigating the optimization of the IRS, we consider the convergence of a realistic system towards asymptotic behavior. We also investigate the use of phase optimization which require minimal channel state information.