### 2020-08-26 documentation/Paper/abstract.tex

Surfaces with reflective phase control (IRS) have been the subject of interest in recent years as means to better control the wireless propagation environment. This work investigates the impact of an IRS on the capacity of MIMO communication systems. Using the tools of free probability, a system is analyzed in the asymptotic domain to derive an expression for the system capacity. A key finding is that in the asymptotic domain, the choice for the phase of reflected waves has no impact for the asymptotic analysis. Motivated the contradiction between this finding and the substantial previous literature investigating the optimization of the IRS, we consider the convergence of a realistic system towards asymptotic behavior. We then investigate new methods for phase optimization for systems without full channel state information at the IRS.