### 2020-09-10 documentation/Optimization Work/09_10_20.tex
Added some details for progress report

\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{graphicx,tikz}
\usepackage{float}
\usetikzlibrary{positioning}
\input defs.tex
\bibliographystyle{ieeetr}
\graphicspath{ {./figures/} }

\title{Progress Report}
\author{Peter Hartig}

\begin{document}
\maketitle
\tableofcontents
\section{System Overview}
The communication system model is given by 
\begin{equation}
\vy = \mH_{Total} \mW \vs  + \vn
\end{equation}
with 
	\begin{equation*}
	\mH_{Total} = \underbrace{\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}}_{\text{IRS}} + \underbrace{\mathbf{G}}_{\text{LOS}}.
	\end{equation*}
	The IRS component of the channel can be decomposed in the paths through the individual elements as
	\begin{equation}
	\mathbf{H}_{2}\boldsymbol{\Phi}\mathbf{H}_{1}= 
	\Sigma_{i=1}^{N_S} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i}
	\end{equation}
	for an IRS with $N_S$ reflective elements. 
In the following we will evaluate the performance of this system using the Mean Square Error given by 
\begin{equation}
E\left[\|  \vs - (\mG + \Sigma_{i=1}^{N_S} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})\mathbf{W}\vs) \|^2 \right]
\end{equation}
in which $\mathbf{W}$ is the linear precoder at the transmitter. Note that receiver noise is not considered in this utility function. 
In the following ,we assume that the elements of the IRS can adjust the phases of the impinging waves and may also be "turned off" such that they reflect any wave.
CSI is estimated using the following procedure.
\begin{enumerate}
\item
We leave all IRS elements \emph{on} and set the phase to a known value (e.g. $\phi_i = 1 $) to first find the full channel   $\mH_{Total}$(and then transmit this to the Tx). At this point, we will not have sufficient information to estimate the individual components of the IRS component of the channel $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$ and thus we cannot optimize with respect to $\phi_i $.
\item
	In the next iteration of pilots, we turn off all elements in order to estimate the LOS matrix $\mG $.
\item
	Next, by turning on a specific IRS element and sending additional pilots, we can estimate 
	the path through a specific IRS element given by $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$.
	Using this information, we can now use the expression 
	\begin{equation}
E\left[\|  \vs - (\mG +  \mH_1 + \Sigma_{i=1}^{1} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})\mathbf{W}\vs  \|^2 \right]
\end{equation}
in which $\mH_{Total} =  \mH_1 + \Sigma_{i=1}^{1} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$
	\item
		Repeating the procedure above, we estimate additional elements $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$ and perform the optimization over increasing number of phases.
	\item 
		We are investigating how this optimization enhances performance w.r.t the MSE as well a how the correlation in the channel may change this. 
		Note that once correlation is considered, the order in which the channels through each element $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$ may impact the optimization performance .
\end{enumerate}
\section{Optimization Problem}
First, we rewrite the above utility function as
	\begin{equation}
E\left[ \trace\left((\vs - \mathbf{H}\mathbf{W}\vs)(\vs - \mathbf{H}\mathbf{W}\vs)^H \right)\right]
\end{equation}
or equivalently
	\begin{equation}
 \trace\left((\mI - \mathbf{H}\mathbf{W})E\left[\vs\vs^H\right](\mI - \mathbf{H}\mathbf{W})^H \right)
\end{equation}
Assuming $\mathbf{H}$ and $\mathbf{W}$ to be constant over a coherence period, the expectation applies only to $\vs \vs^H$. If we consider equal power allocation at the transmitter, this becomes
	\begin{equation}
\trace\left((\mI - \mathbf{H}\mathbf{W})(\mI - \mathbf{H}\mathbf{W})^H \right).
\end{equation}
The MSE optimization with respect to the IRS phases is given by
	\begin{align}
	    \underset{\phase}{\text{minimize }}
	    & \; \trace\left((\mI - \mathbf{H}\mathbf{W})
	   (\mI - \mathbf{H}\mathbf{W})^H \right)
	     \\
	    \text{subject to  } \; &
	    | \phi_i | = 1  \forall i \in {1 \cdots	 K}
	\end{align}\label{mse_problem}
	which is further simplified into
	\begin{subequations}
	\label{mse_problem_s}
	\begin{align}
	   	    \underset{\phi_k}{\text{minimize }}
	    & \; - \trace \left( \mathbf{H}\mathbf{W}\right) - \trace \left( \mathbf{H}\mathbf{W}\right)^* +
	    \trace\left(\mathbf{H}\mathbf{W} \mathbf{W}^H \mathbf{H}^H\right)
	     \\
	    \text{subject to  } \; &
	    | \phi_i | = 1   \; \forall i \in {1 \cdots	 K}
	\end{align}
	\end{subequations}

	with $\mathbf{H} = (\mG +  \mH_1 + \Sigma_{i=1}^{K} \phi_i \mathbf{h}_{2,i}\mathbf{h}^T_{1,i})$ (to fit on page).
	Expanding the above utility function and neglecting the terms without any phase coefficients, the cost function becomes
	\begin{equation}
	\Sigma_{i=1}^{N} \left(\prod_{k=1}^{M_i}\phi_{k} c_i \right ) +  \Sigma_{i=1}^{N} \left(\prod_{k=1}^{M_i}\phi_{k} c_i \right)^H*
	\end{equation}
	From this expression it can be seen that the complex component of each term will be canceled by its complex conjugate.
	One method to perform this optimization is to iterate over each element of the IRS and choose $\phi_{i}$ such that 
	the corresponding component of the sum is real and negative. Each step of the algorithm solves the problem
	\begin{subequations}
	\label{algo_step}
	\begin{align}
	   	    \underset{\phi_k}{\text{minimize }}
	    & \; \phi_k c + \phi_k^* c^* + a 
	     \\
	    \text{subject to  } \; &
	    | \phi_i | = 1   \; \forall i \in {1 \cdots	 K}
	\end{align}
	\end{subequations}
	in which $a \in \reals$ and $c \in \complex$ so the result is simply to choose $\phi_k$ such that $\phi_k c$ is negative and real.
	Note that if the transmitter has found $K \leq N_S$ IRS element channels, $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$, then the optimization can only be performed over
	the elements whose channel , $\mathbf{h}_{2,i}\mathbf{h}^T_{1,i}$, is known (i.e. Step 3 of the pilot symbols has been performed for $\phi_i$).

\section{Optimization Implementation}
	First, I considered the case in which an arbitrary linear precoding filter is used at the transmitter (in this case the matched filter to the channel when all IRS elements have phase $\phi_i = 1 $).
	Each step of the algorithm is the same as Problem \eqref{algo_step}.
	All channel elements, transmit symbols and noise are generated using $\mathcal{CN}(0,1)$.
	The results are shown in Figures \ref{MSE_opt} and \ref{MSE_opt_noise} for the cases with and without noise at the receiver. These results indicate that increasing the number of elements at the IRS can allow for improved optimization gains.
	\begin{figure}[H]
	\includegraphics[width= 16cm,height = 13cm]{figures/MSE_optimized}
	  \caption{MSE with/without optimization versus number of IRS elements without receiver noise.}
	  	  \label{MSE_opt}
	\end{figure}
		The result in Figure \ref{MSE_opt} seems  strange because for the infinite SNR case (i.e. no receiver noise), the matched filter should have optimal MSE right?
	\begin{figure}[H]
	\includegraphics[width= 16cm,height = 13cm]{figures/MSE_optimized_noise}
	  \caption{MSE with/without optimization versus number of IRS elements with receiver noise.}
	  	  \label{MSE_opt_noise}
	\end{figure}

%For certain choices of $\mathbf{W}$ this phase optimization could be performed during the phase optimization step of an alternating optimization algorithm.
\section{Next Steps}
\begin{itemize}
\item
The next step is to extend the MSE/phase optimization to also include the impact of the phases on the beamformer. For example in the case of the matched filter, this would take the form of 
		\begin{align}
	    \underset{\phase}{\text{minimize }}
	    & \; - \trace \left( \mathbf{H}\mathbf{H}\right) - \trace \left( \mathbf{H}\mathbf{H}\right)^* +
	    \trace\left(\mathbf{H}\mathbf{H} \mathbf{H}^H \mathbf{H}^H\right)
	     \\
	    \text{subject to  } \; &
	    | \phi_i | = 1  \forall i \in {1 \cdots	 K}.
	\end{align}\label{mse_problem_full}
To simplify the initial optimization setup in the above results, the beamformer was fixed before the MSE optimization w.r.t. the phases. For a linear beamformer, this should result in a similar optimization setup as \ref{mse_problem_s}, however, finding the coefficients to the phases will be more difficult and there will be many more terms in the sum.  
\item
	Next, the impact of CSI knowledge on the optimization gain should be studied. In this case, only some of the IRS paths should be used in the optimization and the result should be compared to the full CSI case. 
\item
	Finally, the impact of correlation of the impact of the phase optimization should be studied. And perhaps look at the optimal order in which to estimate CSI for the correlated system if not all channels will be found. 
\end{itemize}
\bibliography{bibliography}
\end{document}