### 2020-09-19 documentation/Paper/abstract.tex
updates to report

Surfaces which can control the phase of reflected radio waves, known as intelligent reflective surfaces (IRS), have been the subject of interest in recent years as a means to better control the wireless propagation environment. This work investigates the impact of an IRS on the MIMO communication systems. Using the tools of free probability, a system is analyzed in the asymptotic domain to derive an expression for the system capacity. A key finding is that the phase control offered by the IRS has no impact in the asymptotic domain. Motivated by the contradiction between this finding and the substantial previous literature investigating the optimization of the IRS, we consider the convergence of a realistic system towards asymptotic behavior. We then propose an optimization method for the non-convex problem of minimizing the Mean Square Error with respect to the IRS phases.