### 2020-09-19 documentation/Paper/non_asymptotic.tex

Up to this point we have shown an analysis which seems to contradict the observations from previous investigations reviewed in Section \ref{Background} showing the benefits of exploiting phase control at the IRS. The key to the coexistence of these results is determining when the asymptotic assumptions required by Free Probability hold. 
As mentioned in in Section \ref{Background}, \cite {} have shown that for matrices of random elements, the  convergence of the matrix characteristics towards that predicted by Free Probability is slowed by adding correlations between elements. We would now like to at least numerically investigate this question to investigate the performance of RIS MIMO systems with parameters reflecting state of the art.
(This is where being able to analyze the capacity using the AED from above would provide nice results)
 As seen in \ref{Background} current large systems have been shown for 
sizes S= 224. We now consider whether this size is reasonable under correlations realized in systems (NEED TO FIND and justify).